A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal

### 2020-09-28 Manuscript/2020-MEE-Adams_et_al-Physignal.Rmd
MEE

Tests revealed that across simulation conditions, the average type I error of $\hat{Z}_{12}$ was approximately 0.05 (Fig. 4). Likewise, false discovery rates were also low, and were at or below the nominal 5% level (Fig. 4). Importantly, the statistical performance of $\hat{Z}_{12}$ appears unaffected by phylogeny size. Overall, these results reveal that tests based on $\hat{Z}_{12}$ have acceptable type I error and false discovery rates. Therefore, $\hat{Z}_{12}$ is an appropriate statistic for comparing the degree of phylogenetic signal across traits.

@@

Figure 4. Type I error rates and model misspecification for $\hat{Z}_{12}$. Type I error is found on the far-left of the plot ($\lambda = 0$).

@@

```{r, echo = FALSE, out.height="75%", fig.cap="Type I error rates and model misspecification for $\\hat{Z}_{12}$. Type I error is found on the far-left of the plot ($\\lambda = 0$)."}
include_graphics("Figs/fig.4.png")  
```

>>>>

Tests revealed that across simulation conditions, the average type I error of $\hat{Z}_{12}$ was approximately 0.05 (Fig. 4). Likewise, average false discovery rates were also low, and were at or below the nominal 5% level (Fig. 4). Importantly, the statistical performance of $\hat{Z}_{12}$ appears unaffected by phylogeny size (Supporting Information). Overall, these results reveal that tests based on $\hat{Z}_{12}$ have acceptable type I error and false discovery rates. Therefore, $\hat{Z}_{12}$ is an appropriate statistic for comparing the degree of phylogenetic signal across traits.

@@

Figure 4. Type I error rates and model misspecification for $\hat{Z}_{12}$. Type I error is found on the far-left of the plot ($\lambda = 0$). The black line signifies the average type I error and false discovery rates across simulations of different tree sizes.

@@

```{r, echo = FALSE, out.height="75%", fig.cap="Type I error rates and model misspecification for $\\hat{Z}_{12}$. Type I error is found on the far-left of the plot ($\\lambda = 0$). The black line signifies the average type I error and false discovery rates across simulations of different tree sizes."}
include_graphics("Figs/fig.4.png")  
```

### 2020-09-28 Manuscript/2020-MEE-Adams_et_al-Physignal.tex
MEE

\begin{longtable}[]{@{}rr@{}}
\toprule
words & characters\tabularnewline
\midrule
\endhead
4201 & 26493\tabularnewline
\bottomrule
\end{longtable}

@@

Tests revealed that across simulation conditions, the average type I
error of \(\hat{Z}_{12}\) was approximately 0.05 (Fig. 4). Likewise,
false discovery rates were also low, and were at or below the nominal
5\% level (Fig. 4). Importantly, the statistical performance of
\(\hat{Z}_{12}\) appears unaffected by phylogeny size. Overall, these
results reveal that tests based on \(\hat{Z}_{12}\) have acceptable type
I error and false discovery rates. Therefore, \(\hat{Z}_{12}\) is an
appropriate statistic for comparing the degree of phylogenetic signal
across traits.

@@

Figure 4. Type I error rates and model misspecification for
\(\hat{Z}_{12}\). Type I error is found on the far-left of the plot
(\(\lambda = 0\)).

@@

\begin{figure}
\includegraphics[width=28.51in,height=0.75\textheight]{Figs/fig.4} \caption{Type I error rates and model misspecification for $\hat{Z}_{12}$. Type I error is found on the far-left of the plot ($\lambda = 0$).}\label{fig:unnamed-chunk-6}
\end{figure}

>>>>

\begin{longtable}[]{@{}rr@{}}
\toprule
words & characters\tabularnewline
\midrule
\endhead
4221 & 26607\tabularnewline
\bottomrule
\end{longtable}

@@

Tests revealed that across simulation conditions, the average type I
error of \(\hat{Z}_{12}\) was approximately 0.05 (Fig. 4). Likewise,
average false discovery rates were also low, and were at or below the
nominal 5\% level (Fig. 4). Importantly, the statistical performance of
\(\hat{Z}_{12}\) appears unaffected by phylogeny size (Supporting
Information). Overall, these results reveal that tests based on
\(\hat{Z}_{12}\) have acceptable type I error and false discovery rates.
Therefore, \(\hat{Z}_{12}\) is an appropriate statistic for comparing
the degree of phylogenetic signal across traits.

@@

Figure 4. Type I error rates and model misspecification for
\(\hat{Z}_{12}\). Type I error is found on the far-left of the plot
(\(\lambda = 0\)). The black line signifies the average type I error and
false discovery rates across simulations of different tree sizes.

@@

\begin{figure}
\includegraphics[width=24.22in,height=0.75\textheight]{Figs/fig.4} \caption{Type I error rates and model misspecification for $\hat{Z}_{12}$. Type I error is found on the far-left of the plot ($\lambda = 0$). The black line signifies the average type I error and false discovery rates across simulations of different tree sizes.}\label{fig:unnamed-chunk-6}
\end{figure}

### 2020-09-28 Manuscript/SI.Rmd
MEE

Here we examine the statistical properties of tests for comparing the relative strength of phylogenetic signal, using $\hat{Z}_{12}$. Simulations were performed using pure-birth trees of different sizes ($n=2^5, 2^6, \cdots, 2^{10}$), and with differing levels of phylogenetic signal ($\lambda=0.0, 0.5, \cdots, 1.0$). For each combination of $n$ and $\lambda$, 2 pure-birth trees were generated and 50 traits were simulated on each phylogeny under a Brownian motion model of evolution. Next, the phylogenetic signal between traits was compared in pairwise fashion for all $50*49/2 = 1225$ combinations, using $\hat{Z}_{12}$. The proportion of significant results provided an estimate of the type I error or model misspecification. Specifically, type I error was evaluated when traits were simulated using no input phylogenetic signal (i.e., $\lambda = 0$). Likewise, model misspecification was evaluated for data simulated with some known, but equal, level of phylogenetic signal (i.e., $\lambda > 0$). \hfill\break

**Results:** Tests revealed that across simulation conditions, the average type I error of $\hat{Z}_{12}$ was approximately 0.05 (Fig. S21). Likewise, levels of misspecification were also low, and were at or below the nominal 5% level (Fig. S21 black line). Both results demonstrate that comparisons of the strength of phylogenetic signal using $\hat{Z}_{12}$ do not display high levels of false positives, and are thus appropriate for comparing the degree of phylogenetic signal across traits.

@@

\textbf{Figure S21}. Type I error rates and model misspecification for $\hat{Z}_{12}$. Type I error is found on the far-left of the plot ($\lambda = 0$).

>>>>

Here we examine the statistical properties of tests for comparing the relative strength of phylogenetic signal, using $\hat{Z}_{12}$. Simulations were performed using pure-birth trees of different sizes ($n=2^5, 2^6, \cdots, 2^{10}$), and with differing levels of phylogenetic signal ($\lambda=0.0, 0.5, \cdots, 1.0$). For each combination of $n$ and $\lambda$, 2 pure-birth trees were generated and 50 traits were simulated on each phylogeny under a Brownian motion model of evolution. Next, the phylogenetic signal between traits was compared in pairwise fashion for all $50*49/2 = 1225$ combinations, using $\hat{Z}_{12}$. The proportion of significant results provided an estimate of the type I error or false discovery rate. Specifically, type I error was evaluated when traits were simulated using no input phylogenetic signal (i.e., $\lambda = 0$). Likewise, false discovery rates were evaluated for data simulated with some known, but equal, level of phylogenetic signal (i.e., $\lambda > 0$). \hfill\break

**Results:** Tests revealed that across simulation conditions, the average type I error of $\hat{Z}_{12}$ was approximately 0.05 (Fig. S21). Likewise, levels of false discovery were also low, and were at or below the nominal 5% level (Fig. S21). Both results demonstrate that comparisons of the strength of phylogenetic signal using $\hat{Z}_{12}$ do not display high levels of false positives, and are thus appropriate for comparing the degree of phylogenetic signal across traits.

@@

\textbf{Figure S21}. Type I error rates and false discovery rates for $\hat{Z}_{12}$. Type I error is found on the far-left of the plot ($\lambda = 0$). Each colored line represents the curve for a given level of tree size.

### 2020-09-28 Manuscript/SI.tex
MEE version
MEE

\includegraphics[width=0.95\linewidth]{fig.S1}

@@

\includegraphics[width=0.95\linewidth]{fig.S2}

@@

\includegraphics[width=0.95\linewidth]{fig.S3}

@@

\includegraphics[width=0.95\linewidth]{fig.S4}

@@

\includegraphics[width=0.95\linewidth]{fig.S5}

@@

\includegraphics[width=0.95\linewidth]{fig.S6}

@@

\includegraphics[width=0.95\linewidth]{fig.S7}

@@

\includegraphics[width=0.95\linewidth]{fig.S8}

@@

\includegraphics[width=0.95\linewidth]{fig.S9}

@@

\includegraphics[width=0.95\linewidth]{fig.S10}

@@

\includegraphics[width=0.95\linewidth]{fig.S11}

@@

\includegraphics[width=0.95\linewidth]{fig.S12}

@@

\includegraphics[width=0.95\linewidth]{fig.S13}

@@

\includegraphics[width=0.95\linewidth]{fig.S14}

@@

\includegraphics[width=0.95\linewidth]{fig.S15}

@@

\includegraphics[width=0.95\linewidth]{fig.S16}

@@

\includegraphics[width=0.95\linewidth]{fig.S17}

@@

\includegraphics[width=0.95\linewidth]{fig.S18}

@@

\includegraphics[width=0.95\linewidth]{fig.S19}

@@

\includegraphics[width=0.95\linewidth]{fig.S20}

@@

Here we examine the statistical properties of tests for comparing the
relative strength of phylogenetic signal, using \(\hat{Z}_{12}\).
Simulations were performed using pure-birth trees of different sizes
(\(n=2^5, 2^6, \cdots, 2^{10}\)), and with differing levels of
phylogenetic signal (\(\lambda=0.0, 0.5, \cdots, 1.0\)). For each
combination of \(n\) and \(\lambda\), 2 pure-birth trees were generated
and 50 traits were simulated on each phylogeny under a Brownian motion
model of evolution. Next, the phylogenetic signal between traits was
compared in pairwise fashion for all \(50*49/2 = 1225\) combinations,
using \(\hat{Z}_{12}\). The proportion of significant results provided
an estimate of the type I error or model misspecification. Specifically,
type I error was evaluated when traits were simulated using no input
phylogenetic signal (i.e., \(\lambda = 0\)). Likewise, model
misspecification was evaluated for data simulated with some known, but
equal, level of phylogenetic signal (i.e., \(\lambda > 0\)).
\hfill\break

\textbf{Results:} Tests revealed that across simulation conditions, the
average type I error of \(\hat{Z}_{12}\) was approximately 0.05 (Fig.
S21). Likewise, levels of misspecification were also low, and were at or
below the nominal 5\% level (Fig. S21 black line). Both results
demonstrate that comparisons of the strength of phylogenetic signal
using \(\hat{Z}_{12}\) do not display high levels of false positives,
and are thus appropriate for comparing the degree of phylogenetic signal
across traits.

\includegraphics[width=6.67in,height=0.5\textheight]{fig.S21}

\textbf{Figure S21}. Type I error rates and model misspecification for
\(\hat{Z}_{12}\). Type I error is found on the far-left of the plot
(\(\lambda = 0\)). The solid black line represents the average type I
error and false discovery rate across all levels of input signal.

>>>>

\includegraphics[width=0.95\linewidth]{Figs/fig.S1}

@@

\includegraphics[width=0.95\linewidth]{Figs/fig.S2}

@@

\includegraphics[width=0.95\linewidth]{Figs/fig.S3}

@@

\includegraphics[width=0.95\linewidth]{Figs/fig.S4}

@@

\includegraphics[width=0.95\linewidth]{Figs/fig.S5}

@@

\includegraphics[width=0.95\linewidth]{Figs/fig.S6}

@@

\includegraphics[width=0.95\linewidth]{Figs/fig.S7}

@@

\includegraphics[width=0.95\linewidth]{Figs/fig.S8}

@@

\includegraphics[width=0.95\linewidth]{Figs/fig.S9}

@@

\includegraphics[width=0.95\linewidth]{Figs/fig.S10}

@@

\includegraphics[width=0.95\linewidth]{Figs/fig.S11}

@@

\includegraphics[width=0.95\linewidth]{Figs/fig.S12}

@@

\includegraphics[width=0.95\linewidth]{Figs/fig.S13}

@@

\includegraphics[width=0.95\linewidth]{Figs/fig.S14}

@@

\includegraphics[width=0.95\linewidth]{Figs/fig.S15}

@@

\includegraphics[width=0.95\linewidth]{Figs/fig.S16}

@@

\includegraphics[width=0.95\linewidth]{Figs/fig.S17}

@@

\includegraphics[width=0.95\linewidth]{Figs/fig.S18}

@@

\includegraphics[width=0.95\linewidth]{Figs/fig.S19}

@@

\includegraphics[width=0.95\linewidth]{Figs/fig.S20}

@@

Here we examine the statistical properties of tests for comparing the
relative strength of phylogenetic signal, using \(\hat{Z}_{12}\).
Simulations were performed using pure-birth trees of different sizes
(\(n=2^5, 2^6, \cdots, 2^{10}\)), and with differing levels of
phylogenetic signal (\(\lambda=0.0, 0.5, \cdots, 1.0\)). For each
combination of \(n\) and \(\lambda\), 2 pure-birth trees were generated
and 50 traits were simulated on each phylogeny under a Brownian motion
model of evolution. Next, the phylogenetic signal between traits was
compared in pairwise fashion for all \(50*49/2 = 1225\) combinations,
using \(\hat{Z}_{12}\). The proportion of significant results provided
an estimate of the type I error or false discovery rate. Specifically,
type I error was evaluated when traits were simulated using no input
phylogenetic signal (i.e., \(\lambda = 0\)). Likewise, false discovery
rates were evaluated for data simulated with some known, but equal,
level of phylogenetic signal (i.e., \(\lambda > 0\)). \hfill\break

\textbf{Results:} Tests revealed that across simulation conditions, the
average type I error of \(\hat{Z}_{12}\) was approximately 0.05 (Fig.
S21). Likewise, levels of false discovery were also low, and were at or
below the nominal 5\% level (Fig. S21). Both results demonstrate that
comparisons of the strength of phylogenetic signal using
\(\hat{Z}_{12}\) do not display high levels of false positives, and are
thus appropriate for comparing the degree of phylogenetic signal across
traits.

\includegraphics[width=6.67in,height=0.5\textheight]{Figs/fig.S21}

\textbf{Figure S21}. Type I error rates and false discovery rates for
\(\hat{Z}_{12}\). Type I error is found on the far-left of the plot
(\(\lambda = 0\)). Each colored line represents the curve for a given
level of tree size.
